---
title: "Data Pull User Guide"
author: "Tetra Tech"
format: docx
editor: visual
---

## Required Packages

The required packages for the data_pull.R code are *TADA* and *tidyverse*. In order to install the *TADA* package, you will also need the *remotes* package. The *remotes* package is not required otherwise. Here are the lines to install these packages:

```{r}
#Install tidyverse
install.packages('tidyverse')

#Install remotes
install.packages("remotes",
  repos = "http://cran.us.r-project.org"
)
library(remotes)

#Install TADA
remotes::install_github("USEPA/TADA",
  ref = "develop",
  dependencies = TRUE
)

```

## Define Date Parameters

The only necessary input for the data pull is the start and end dates. Change the dates using the format 'YYYY-MM-DD' for startDate and endDate for whatever is required.

```{r}
startDate <- '2018-02-01'
endDate <- '2023-02-01'
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
